Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-05-02 SupportingInformation.Rmd
v1.description text moved from SupportingInformation.Rmd to Paper.Rmd

\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{section}{0}
\renewcommand{\thesection}{S\arabic{section}}

@@

<!--New latex commands-->
\newcommand{\mean}{\operatorname{mean}}
\newcommand{\var}{\operatorname{var}}
\newcommand{\cor}{\operatorname{cor}}
\newcommand{\Ps}{\operatorname{P}}  
\newcommand{\Dsq}{\operatorname{D}^2}
\newcommand{\rank}{\operatorname{rank}}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/SoilCNraw.pdf}
\includegraphics[width=3 in]{./Results/SoilCNcop.pdf}
\caption{\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data. Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n = `r nrow(d)` $ is the number of sampling locations (see Sec. \ref{Model_selection}).\label{fig_raw_cop_soilCN}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/BMR_results/rawdataplot_BMR_vs_lnmass.pdf}
\includegraphics[width=3 in]{./Results/BMR_results/copula_BMR_vs_lnmass.pdf}
\caption{\textbf{(A)} Plot of the raw BMR vs. ln(mass) data \textbf{(B)} Copula plot for BMR dataset. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see Sec. \ref{Model_selection}).\label{fig_raw_cop_BMR}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
\caption{\textbf{(A)} Plot of the raw AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. \textbf{(B)} Copula plot of AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see Sec. \ref{Model_selection}). \label{fig_raw_cop_AVG_humerus_ltoe}}
\end{center}
\end{figure}

@@

![Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas. See text for details. $n=35$. \label{fig_stat_testing35}](./Results/stat_results/stat_testing/res_pt35_rankT.pdf)

![Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas. See text for details. $n=1000$.\label{fig_stat_testing1000}](./Results/stat_results/stat_testing/res_pt1000_rankT.pdf)

# Question 1 analyses for each dataset
Recall that question 1 is: Do datasets in ecology and related fields have non-Gausian copula structure? Do they show tail dependence distinct from that of a Gaussian copula? In particular do they show asymmetric tail dependence?

<!--
## Details of methods \label{Method}

### Model selection approach \label{Model_selection}

### Non-parametric analysis of tail dependence \label{NPA}

moved to Paper.Rmd
-->

<!--
## Results: soil C and N data \label{Ex_SoilCN}

### Model selection approach \label{Ex_SoilCN_Model_selection}

moved to Paper.Rmd
-->

<!--This does the computations for this section-->
```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
families_Loecke<-c(1,3:10,13:14,16:20) 
BivMS_res_Loecke<-OurBiCopSelect(u1=v_CN[,1],u2=v_CN[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
saveRDS(BivMS_res_Loecke,file="./Results/fitting_results/BivMS_soilCN.RDS")
```

@@

### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}

```{r stat_soilCN_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_CN,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
stat_soilCN_ga<-bivfunctionplot(v=v_CN,resloc="./Results/stat_results/stat_soilCN/",nametag="soilCN",numbin=10)
```
<!--****DAN: DAN suspects a bug in the code that generates this fig. because the red dot should appear outside the x's iff the text > 999 (or so,mething like that) appears in the bin, also check make_table_stat_fn r chunk-->

As described in Methods, each of our statistics, as computed for the 
ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$, was compared to the distributions 
of values of the same statistic with the same ranges computed on 1000 normal-copula 
surrogate datasets with similar Kendall or Spearman correlations, in different 
runs for Kendall and Spearman (Figure \ref{fig_soilCN_nonparam}). Because the correlation 
used in the $\cor$ statistic is the Spearman correlation, Spearman-preserving surrogates
should be used, so the top-right panel of Figure \ref{fig_soilCN_nonparam}
is more reliable than the top left. Results indicate that tail dependence of 
data is stronger in both the lower and upper tails than would be expected under a 
null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
weaker in the middle of the distributions
than would be expected based on a normal-copula null model.

![Nonparametric tests for tail dependence and other deviations from a normal copula for soil C and N data. As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).  When red lines were outside dashed lines, text at the top of plots indicates how many surrogate values the data value was less than or greater than. When $\cor$ values (respectively, $\Ps$ or $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, stronger than or weaker than for $\Ps$ or $\Dsq$) expected from a normal-copula null hypothesis. \label{fig_soilCN_nonparam}](./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf)

We also tested statistically for asymmetry of tail dependence, which seems visible in Figure \ref{fig_raw_cop_soilCN}(B). 
Results (Table \ref{tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.

@@

## Results: BMR data \label{Ex_BMR}

### Model selection approach \label{Ex_BMR_Model_selection}

```{r bivMS_BMR, echo=F, cache=T, cache.extra=list(seed,v_BMR,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
set.seed(seed)
source("OurBiCopSelect.R")
#families<-c(1,3:10,13:14,16:20) 
BivMS_BMR<-OurBiCopSelect(u1=v_BMR[,1],u2=v_BMR[,2],families=c(1,3:10,13:14,16:20),
                          level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
saveRDS(BivMS_BMR,file="./Results/BMR_results/BivMS_BMR_lnmass.RDS")
```

The test of indepedence gave $p=$ `r BivMS_BMR$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{tab_BMRfit}). The table shows that Frank is much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

@@

The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was $0$ as expected, because the contributions came from Frank and SBB8 which are both symmetric copula.
<!--$`r BivMS_BMR$relLTdep_AICw`$ and the model-averaged upper-tail dependence statistic was 
$`r BivMS_BMR$relUTdep_AICw`$. These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), 
$`r BivMS_BMR$relLTdep_AICw-BivMS_BMR$relUTdep_AICw`$, was different from 0, answering the third part of question 1. 
-->
We also tested that lowest-AIC copula Frank was a good-enough fit, giving $p$-value 
$`r BivMS_BMR$GofRes_CvM`$ (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value $`r BivMS_BMR$GofRes_KS`$ in the Kolmogorov-Smirnov-based test of the goodness of fit. The next section looks at tail dependence in a non-parametric way.

### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}

```{r stat_BMR_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_BMR,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
# Now do NPA for v_BMR
# If "F"/"SBB8" is the best fit : then it should be symmetric tail dep (0) : Then still I need to do NPA ?
# May be : because it will also tell me how they are different from "N" copula in different sections?
source("bivfunctionplot.R")
stat_BMR_ga<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_10",numbin = 10)
stat_BMR_ga_numbin5<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_5",numbin = 5)
```

![Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data. As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles). \label{fig_BMR_nonparam}](./Results/BMR_results/BMR_numbin_10_bivfunctionplot.pdf)

![Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data. As described in Methods, each statistic was computed on the data using the ranges $0-0.2$, $0.2-0.4$, ..., $0.8-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles). \label{fig_BMR_nonparam5}](./Results/BMR_results/BMR_numbin_5_bivfunctionplot.pdf)

@@

## Results: contrasts.csv data \label{Ex_contrasts}

### Model selection approach \label{Ex_contrasts_Model_selection}

```{r bivMS_contrasts_67, echo=F, cache=T, cache.extra=list(seed,v_contrasts_67,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
set.seed(seed)
# Now do Model Selection for [i,j]=[6,7]
# 6 : AVG. ltoe.IV
# 7 : AVG. humerus
source("OurBiCopSelect.R")
families<-c(1,3:10,13:14,16:20)
BivMS_contrasts_67<-OurBiCopSelect(u1=v_contrasts_67[,1],u2=v_contrasts_67[,2],families=families,
                          level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                          gofnormal=F,status=F)
saveRDS(BivMS_contrasts_67,file="./Results/Anolis_results/contrasts_csv_results/BivMS_contrasts_67.RDS")
```

The test of indepedence gave $p=$ `r BivMS_contrasts_67$IndepTestRes` (to within the precision available 
from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{tab_contrasts_67fit}). The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

@@

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`. These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1.

These results do convincingly show non-Gaussian copula structure, giving $p$-value `r BivMS_contrasts_67$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_contrasts_67$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. The next section looks at tail dependence in a non-parametric way.

### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}

```{r stat_contrasts_67_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_contrasts_67,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
# Now do NPA for v_contrasts_67
stat_contrasts_67_ga<-bivfunctionplot(v=v_contrasts_67,resloc="./Results/Anolis_results/contrasts_csv_results/",
                                      nametag="contrasts_67",numbin = 4)
```

![Nonparametric tests for tail dependence and other deviations from a normal copula for contrasts_67 data. As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles). \label{fig_BMR_nonparam}](./Results/Anolis_results/contrasts_csv_results/contrasts_67_bivfunctionplot.pdf)

@@

## Results: abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}

### Model selection approach \label{Ex_count_Model_selection}

@@

<!--Do the calculations-->
```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}

@@

As described in Methods, model selection among copulas was done separately for each pair of the 10 
sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns 
corresponding to locations. Table \ref{tab_count_bestcop} shows that, for the large majority of location 
pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of 
question 1 for these data. Comparing the 
AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{tab_count_AICw}) 
to the AIC weight 
of the normal copula family for the same pair (Table \ref{tab_count_AICw_normal}) confirms that the degree of support for normal copulas was usually not
comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{tab_count_p_KS}),
demonstrating that the collection of copula families we used was sufficiently broad that the copulas of 
all pairs of locations could be characterized by at least one of them. In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

```{r read_RES_aphid_count, echo=F}
RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
```

@@

The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{tab_count_LTdep_AICw} and \ref{tab_count_UTdep_AICw} respectively. These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data. Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.

@@

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1.

@@

### Non-parametric analysis of tail dependence \label{Ex_count_NPA}

<!--Do the main computations for this section-->
```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

@@

```{r read_stat_aphid_count,echo=F}
stat_aphid_count<-readRDS("./Results/stat_results/stat_aphid_count/stat_aphid_count_sp_10.RDS")
```

Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
resampling scheme, as described in Methods (Table \ref{tab_count_resamp}). Confidence intervals of
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
structure do not have, providing further evidence answering the third part of question 1.

@@

```{r read_RES_aphid_ff,echo=F}
RES_aphid_ff<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_firstflight_species_11.RDS")
```

>>>>

@@

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/SoilCNraw.pdf}
\includegraphics[width=3 in]{./Results/SoilCNcop.pdf}
\caption[\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data]{\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data. Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n = `r nrow(d)` $ is the number of sampling locations (see Sec. \ref{Model_selection}).\label{fig_raw_cop_soilCN}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/BMR_results/rawdataplot_BMR_vs_lnmass.pdf}
\includegraphics[width=3 in]{./Results/BMR_results/copula_BMR_vs_lnmass.pdf}
\caption[\textbf{(A)} Plot of the raw BMR vs. ln(mass) data \textbf{(B)} Copula plot for BMR dataset]{\textbf{(A)} Plot of the raw BMR vs. ln(mass) data \textbf{(B)} Copula plot for BMR dataset. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see Sec. \ref{Model_selection}).\label{fig_raw_cop_BMR}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
\caption[\textbf{(A)} Plot of the raw AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. \textbf{(B)} Copula plot of AVG.ltoe.IV vs. AVG.humerus phylogenetic traits]{\textbf{(A)} Plot of the raw AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. \textbf{(B)} Copula plot of AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see Sec. \ref{Model_selection}). \label{fig_raw_cop_AVG_humerus_ltoe}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt35_rankT.pdf}
\caption[Testing non-parametric statistics with datapoints $n=35$]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas. See text for details. $n=35$. \label{fig_stat_testing35}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt1000_rankT.pdf}
\caption[Testing non-parametric statistics with datapoints $n=1000$]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas. See text for details. $n=1000$.\label{fig_stat_testing1000}}
\end{center}
\end{figure}

<!--
# Question 1 analyses for each dataset
Recall that question 1 is: Do datasets in ecology and related fields have non-Gausian copula structure? Do they show tail dependence distinct from that of a Gaussian copula? In particular do they show asymmetric tail dependence?

## Details of methods \label{Method}
### Model selection approach \label{Model_selection}
### Non-parametric analysis of tail dependence \label{NPA}
moved to Paper.Rmd
-->

<!--
## Results: soil C and N data \label{Ex_SoilCN}
### Model selection approach \label{Ex_SoilCN_Model_selection} 
### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}
                    [[[[moved to Paper.Rmd]]]]
-->

<!--This chunk is for model selection results of soilCN dataset-->
```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
families_Loecke<-c(1,3:10,13:14,16:20) 
BivMS_res_Loecke<-OurBiCopSelect(u1=v_CN[,1],u2=v_CN[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
saveRDS(BivMS_res_Loecke,file="./Results/fitting_results/BivMS_soilCN.RDS")
```

@@

<!--This chunk is for npa testing results of soilCN dataset-->
```{r stat_soilCN_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_CN,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
stat_soilCN_ga<-bivfunctionplot(v=v_CN,resloc="./Results/stat_results/stat_soilCN/",nametag="soilCN",numbin=10)
```
<!--****DAN: DAN suspects a bug in the code that generates this fig. because the red dot should appear outside the x's iff the text > 999 (or so,mething like that) appears in the bin, also check make_table_stat_fn r chunk-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
\caption[Testing non-parametric statistics for soilCN data]{Nonparametric tests for tail dependence and other deviations from a normal copula for soil C and N data. As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).  When red lines were outside dashed lines, text at the top of plots indicates how many surrogate values the data value was less than or greater than. When $\cor$ values (respectively, $\Ps$ or $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, stronger than or weaker than for $\Ps$ or $\Dsq$) expected from a normal-copula null hypothesis. \label{fig_soilCN_nonparam}}
\end{center}
\end{figure}

@@

<!--
\newpage
## Results: BMR data \label{Ex_BMR}
### Model selection approach \label{Ex_BMR_Model_selection}
### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}
                  ----moved to Paper.Rmd----
-->

<!--This chunk is for model selection results of BMR data-->
```{r bivMS_BMR, echo=F, cache=T, cache.extra=list(seed,v_BMR,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
set.seed(seed)
source("OurBiCopSelect.R")
#families<-c(1,3:10,13:14,16:20) 
BivMS_BMR<-OurBiCopSelect(u1=v_BMR[,1],u2=v_BMR[,2],families=c(1,3:10,13:14,16:20),
                          level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
saveRDS(BivMS_BMR,file="./Results/BMR_results/BivMS_BMR_lnmass.RDS")
```

@@

<!--This chunk is for npa testing results of BMR dataset-->
```{r stat_BMR_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_BMR,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
# Now do NPA for v_BMR
# If "F"/"SBB8" is the best fit : then it should be symmetric tail dep (0) : Then still I need to do NPA ?
# May be : because it will also tell me how they are different from "N" copula in different sections?
source("bivfunctionplot.R")
stat_BMR_ga<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_10",numbin = 10)
stat_BMR_ga_numbin5<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_5",numbin = 5)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/BMR_numbin_10_bivfunctionplot.pdf}
\caption[Testing non-parametric statistics for BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data. As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles). \label{fig_BMR_nonparam}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/BMR_numbin_5_bivfunctionplot.pdf}
\caption[Testing non-parametric statistics for BMR data with $numbin=5$]{Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data. As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.2$, $0.2-0.4$, ..., $0.8-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles). \label{fig_BMR_nonparam5}}
\end{center}
\end{figure}

@@

<!--
\newpage
## Results: contrasts.csv data \label{Ex_contrasts}
### Model selection approach \label{Ex_contrasts_Model_selection}
### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}
                 ----moved to Paper.Rmd----
-->

<!--This chunk is for model selection results of anolis data-->
```{r bivMS_contrasts_67, echo=F, cache=T, cache.extra=list(seed,v_contrasts_67,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
set.seed(seed)
# Now do Model Selection for [i,j]=[6,7]
# 6 : AVG. ltoe.IV
# 7 : AVG. humerus
source("OurBiCopSelect.R")
families<-c(1,3:10,13:14,16:20)
BivMS_contrasts_67<-OurBiCopSelect(u1=v_contrasts_67[,1],u2=v_contrasts_67[,2],families=families,
                          level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                          gofnormal=F,status=F)
saveRDS(BivMS_contrasts_67,file="./Results/Anolis_results/contrasts_csv_results/BivMS_contrasts_67.RDS")
```

@@

<!--This chunk is for npa testing results of anolis dataset-->
```{r stat_contrasts_67_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_contrasts_67,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
# Now do NPA for v_contrasts_67
stat_contrasts_67_ga<-bivfunctionplot(v=v_contrasts_67,resloc="./Results/Anolis_results/contrasts_csv_results/",
                                      nametag="contrasts_67",numbin = 4)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Anolis_results/contrasts_csv_results/contrasts_67_bivfunctionplot.pdf}
\caption[Testing non-parametric statistics for anolis data]{Nonparametric tests for tail dependence and other deviations from a normal copula for anolis data. As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles). \label{fig_contrasts_67_nonparam}}
\end{center}
\end{figure}

<!--
{Nonparametric tests for tail dependence and other deviations from a normal copula for anolis data. As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for $1000$ normal-copula surrogate datasets (dashed lines show $0.025$ and $0.975$ quantiles of these results and dotted lines show $0.005$ and $0.995$ quantiles). \label{fig_contrasts_67_nonparam}}
\end{center}
\end{figure}
-->

@@

<!--
\newpage
## Results: abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
### Model selection approach \label{Ex_count_Model_selection}
### Non-parametric analysis of tail dependence \label{Ex_count_NPA}
                  ----moved to Paper.Rmd----
-->

@@

<!--Do the model selection for aphid count data and then showing results in Tables-->
```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}

@@

@@

@@

@@

<!--Do the main computations for non-parametric analysis of aphid count data-->
```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

@@

@@
